Learning STRIPS action models with classical planning

### 2017-10-28 doc/icaps-18/strips-learning.tex
Eva's changes

\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\'ecnica de Val\'encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}

@@

\section{Related work}

The  LIVE  system  (Shen  and  Simon,  1989)  was  an  extension  of  the  General  Problem  Solver  (GPS) framework (Ernst and Newell, 1969) with a learning component. LIVE alternated problem solving with model  learning  to  automatically  define  operators.  The  decision  about  when  to  alternate  depended  onsurprises,  that  is  situations  where  an  action  effects  violated  its  predicted  model.  EXPO  (Gil,  1992) generated plans with the  PRODIGY  system (Minton, 1988), monitored the plans execution, detected differences in the predicted and the observed states and constructed a set of specific hypotheses to fix those differences. Then the  EXPO  filtered the hypotheses heuristically.  OBSERVER  (Wang, 1994) learned operators by monitoring expert agents and applying the version spaces algorithm (Mitchell, 1997) to the observations. When the system already had an operator representation, the preconditions were updated by removing facts that were not present in the new observation’s pre-state; the effects were augmented by adding facts that were in the observation’s delta-state. All of these early works were based on direct liftings of the observed states. They also benefit from experience beyond simple interaction with the environment such as exploratory plans or external teachers, but  none  provided  a  theoretical  justification  for  this  second  source  of  knowledge.  The  work  recently reported in (Walsh and Littman, 2008) succeeds in bounding the number of interactions the learner must complete to learn the preconditions and effects of a STRIPS action model. This work shows that learning STRIPS  operators  from  pure  interaction  with  the  environment,  can  require  an  exponential  number  of samples, but that limiting the size of the precondition lists enable sample-efficient learning (polynomial in the number of actions and predicates of the domain). The work also proves that efficient learning is also possible without this limit if an agent has access to an external teacher that can provide solution traces on demand.

Others systems have tried to learn more expressive action models for deterministic planning in fully observable environments. Examples would include the learning of conditional costs for AP actions (Jess Lanchas and Borrajo, 2007) or the learning of conditional effects with quantifiers (Zhuo et al., 2008).

In addition action model learning has been studied in domains where there is partial state observability. ARMS uses the same kind od learning examples but assumes the exampls are given and proceeds in two phases. In the first phase, ARMS extracts frequent action sets from plans that share a common set of parameters. ARMS also finds some frequent literal-action pairs with the help of the initial state and the goal state that provide an initial guess on the actions preconditions, and effects. In the second phase, ARMS uses the frequent action sets and literal-action pairs to define a set of weighted constraints that must hold in order to make the plans correct. Then, ARMS solves the resulting weighted MAX-SAT problem and produces action models from the solution of the SAT problem. This process iterates until all actions are modelled. For a complex planning domain that involves hundreds of literals and actions, the corresponding weighted MAX-SAT representation is likely to be too large to be solved efficiently as the number of clauses can reach up to tens of thousands. For that reason ARMS implements a hill-climbing method that models the actions approximately. Consequently, the ARMS output is a model which may be inconsistent with the examples.

(Amir and Chang, 2008) introduced an algorithm that tractably generates all the STRIPS-like models that  could  have  lead  to  a  set  of  observations.  Given  a  formula  representing  the  initial  belief  state,  a sequence of executed actions and the corresponding observed states(where partial observations of states are given), it builds a complete explanation of observations by models of actions through a Conjunctive Normal Form (CNF) formula. By linking the possible states of fluents to the effect propositions in the action models, the complexity of the CNF encoding can be controlled to find  exact  solutions  efficiently  in  some  circumstances.  The  learning  algorithm  updates  the  formula  of the belief state with every action and observation in the sequence. This update makes sure that the new formula represents all the transition relations consistent with the actions and observations. The formula returned at the end includes all consistent models, which can then be retrieved with additional processing. Unlike the previous approaches, the one described in (Mour ao et al., 2008) deals with both missing and noisy predicates in the observations. For each action in a given domain, they use kernel perceptrons to learn predictions of the domain properties that change because of the action execution.

In the {\sf LOCM} approach for learning planning models \cite{cresswell2013acquiring}, the only required input are the example training plans, without the need for providing the initial and goal state information. {\sf LOCM} exploits assumptions about the kind of domain model it has to generate. Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}), where each object has a defined set of states that can be captured by a parameterized FSM. The induction of FSM draws upon assumptions like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence. The intuitive assumptions of {\sf LOCM} yield a learning model heavily reliant on the kind of domain structure and not capable of properly deriving domain theories in which the state of a sort is subject to different state machines. This limitation is later overcome by {\sf LOCM2} by forming separate state machines for a sort, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.

>>>>

\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\`ecnica de Val\`encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}

@@

\section{Past and future of learning action models}

Back in the 90's various systems were aimed at learning operators mostly via interaction with the environment. The proposal of the LIVE system was to capture and formulate observable features of objects and use them to acquire and refine operators \cite{ShenS89}. OBSERVER, on the other hand, updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}. These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge. The recent work reported in \cite{WalshL08} shows that although learning STRIPS operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.

Whilst the aforementioned works deal with full state observability, action model learning has also been studied in domains
where there is partial or missing state observability~\cite{jimenez2012review}. The {\sf ARMS} approach~\cite{yang2007learning}, which works when no or partial intermediate states are given, extracts frequent action sets from plans that share a common set of parameters and frequent literal-action pairs with the help of the initial and goal state of the plans. The extracted information is used then to define a set of weighted constraints that must hold for all the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver. Since complex planning problems may result in a too large MAX-SAT representation to be solved efficiently, {\sf ARMS} implements a hill-climbing method that models the actions approximately. Consequently, the output model of {\sf ARMS} may be inconsistent with the examples. The approach presented in \cite{amir:alearning:JAIR08} deals with partial observability. Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation
of observations by models of actions through a CNF formula. The learning algorithm updates the formula of the belief state with every action and observation in the sequence. This update makes sure that the new formula represents all the transition relations consistent with the actions and observations. The formula returned at the end includes all consistent models, which can then be retrieved with additional processing. Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations. An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then STRIPS rules are derived from the classifiers' parameters.

The approach for learning planing models that most likely works with the least information possible is {\sf LOCM} ~\cite{cresswell2013acquiring}, where the only required input are the example training plans without the need for providing the initial and goal state information. {\sf LOCM} exploits assumptions about the kind of domain model it has to generate. Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}) where each object has a defined set of states that can be captured by a parameterized Finite State Machine (FSM). The induction of FSM draws upon assumptions like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence. The intuitive assumptions of {\sf LOCM} yield a learning model heavily reliant on the kind of domain structure and not capable of properly deriving domain theories in which the state of a sort is subject to different state machines. This limitation is later overcome by {\sf LOCM2} by forming separate state machines for a sort, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}. The last contribution of the {\sf LOCM} family, {\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{GregoryC16}), addresses the problem of inducing static predicates in action models by identifying the set of minimal static relations associated to an operator. Because {\sf LOCM} models induce similar models for domains with similar structure (e.g. Blocksworld and Freecell), they face problems at identifying models that do not have static relations (Blocksworld) and those that do have static predicates (Freecell). In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.

==============================================
